A Switching Planner for Combined Task and Observation Planning

### 2010-11-27 code/subarchitectures/planner.sa/trunk/src/latex/ICAPS-2011/switching.tex
typo that distroyed the meanin of a sentence in the switching secion.

Our approach always begins with a sequential session. Non-assumptive
actions from the trace are executed in sequence until the
applicability of the next scheduled action is uncertain. We denote
that action \switchAction.  A contingent session then begins, which
tailors sensory processing by the agent to determine whether the
assumptions made in the trace hold. For assumptions in the trace, we
add a {\em disconfirm} action to the POMDP whose execution is
rewarding (resp. costly) if the assumption is false (resp. true). We
also add one {\em confirm} action that is rewarding (resp. costly) to
execute if $\poss(\switchAction)$ is true (resp. false). If execution
of the contingent plan applies a disconfirm action, then a new
sequential session begins at the {\em underlying} belief-state. If the
confirm action is executed, the sequential session is resumed, and
\switchAction\ is executed. Otherwise, control rests in the continual
session.
%%
Because contingent planning is only practical in relatively small
POMDPs, contingent sessions plan in an abstract decision process
determined by the current trace. This abstraction is constructed by
first omitting all propositions that do not featured in the trace, and
by then iteratively refining the model while the result is of a
practicable size.

Finally, whether proceeding in a sequential or contingent session, our
continual planner maintains a factored representation of successive
belief-states, by performing belief revision according to
Eq~\ref{eq:revision}. Moreover, our internal representation the
underlying belief-state corresponds more-or-less directly to an
$(\pp{:init})$ declaration in DTPDDL. Our approach uses that
distribution: (1) as the source of candidate determinisations for
sequential planning, (2) to determine when to switch, and (3) as a
mechanism to guide construction of an abstract process for contingent
sessions.

@@

\noindent An assumption corresponds to a {\em
relaxed} visitation of the root term of $(\pp{:init})$. In this
relaxed case, a conjunctive term is visited iff its atomic subterms
are visited, and zero or more of its immediate probabilistic subterms
are visited. The starting state, $\state_0$, for sequential planning
is an {\em abstract} state that captures the unique assumption which
has probability $1.0$. In $\pp{DORA}$, that is:

@@

In order to make assumptions available to the sequential planner, we
add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
each element, $\prob_i (T_i)$, of each probabilistic term from
$(\pp{:init})$. The physics of these actions is as
follows. $\assumptiveS{i}$ can be executed if no $\assumptiveS{j}$,
$j \neq i$, has been executed from the same probabilistic term, and,
either $(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root conjunct,
or it occurs in $T_k$ for some executed $\assumptiveS{k}$.
%%
Executing $\assumptiveS{i}$ in $\state$ effects a transition to a
successor state $\state^{T_i}$ with probability $\prob_i$, and
$\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
the union of $\state$ with atomic terms from $T_i$ and conflicting
propositions from $s$ deleted. For example, if the action adds
$(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must delete
$(=(\pp{is-in}~\pp{box})~\#)$. State $\state^\bot$ is an added sink,
where $\forall\action\in\actions$ $\reward(\state^\bot, \action) = 0$
and $\transProb(\state^\bot, \action, \state^\bot) = 1$.

For sequential planning the operators from the POMDP model are
available with semantics that accommodate the above abstraction as
follows. In the deterministic model a proposition $\prop$ can be
thought to have a ternary interpretation at a state \state, as either
{\em true}, written $\prop\in\state$, {\em false}, written
$\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
notation written $\prop\#\in\state$. For example, in $\pp{DORA}$
$(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
$\action\in\actions$ in the deterministic model, if
$\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
or negative effect of $\action$, then $\action$ is not applicable
in \state\ if $\prop\#\in\state$. In other words, an assumptive action
must be applied to make a truth assignment to \prop, and then that
truth assignment could be modified by the DTPDDL domain operators.
%%
Lastly, With regards to switching, we have that if $\action$ has a
precondition that is not true with probability $1$ in the underlying
belief $\bstate$, then immediately before it is executed blindly, the
system switches to a contingent session, discussed in the following
section.

@@

\noindent which is equal to the maximal 
additive contribution a trace of the plan can make to
Eq~\ref{eq:expectedvalue}.  In goal directed problems, where the only
non-zero reward is received at the first transition to a goal state,
the Eq~\ref{eq:tracevalue} criteria gives us the behaviour
of \system{FFR$_a$}~\cite{yoon:etal:2007}. From the perspective of
gradient-based reinforcement learning systems, such as William's
REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
under reasonable assumptions the optimal trace identifies a maximal
gradient step from a uniformly random soft-max policy. Finally, it is
worth clarifying that although the above discussion has been in terms
of a trace of length $N$, there is nothing in our approach that
artificially limits the length of candidate sequential plans.

>>>>

Our approach always begins with a sequential session. Non-assumptive
actions from the trace are executed in sequence until the
applicability of the next scheduled action is uncertain. We denote
that action \switchAction.  A contingent session then begins, which
tailors sensory processing by the agent to determine whether the
assumptions made in the trace hold. For each assumption we add a {\em
disconfirm} action to the POMDP whose execution is rewarding
(resp. costly) if the assumption is false (resp. true). We also add
one {\em confirm} action that is rewarding (resp. costly) to execute
if $\poss(\switchAction)$ is true (resp. false). If execution of the
contingent plan applies a disconfirm action, then a new sequential
session begins at the {\em underlying} belief-state. If the confirm
action is executed, the sequential session is resumed, and
\switchAction\ is executed. Otherwise, control rests in the continual
session.
%%
Because contingent planning is only practical in relatively small
POMDPs, contingent sessions plan in an abstract decision process
determined by the current trace. This abstraction is constructed by
first omitting all propositions that do not featured in the trace, and
by then iteratively refining the model while the result is of a
practicable size.

Finally, whether proceeding in a sequential or contingent session, our
continual planner maintains a factored representation of successive
belief-states, by performing belief revision according to
Eq~\ref{eq:revision}. Moreover, our internal representation of the
underlying belief-state corresponds more-or-less directly to an
$(\pp{:init})$ declaration in DTPDDL. Our approach uses that
distribution: (1) as the source of candidate determinisations for
sequential planning, (2) to determine when to switch, and (3) as a
mechanism to guide construction of an abstract process for contingent
sessions.

@@

\noindent An assumption corresponds to a {\em
relaxed} visitation of the root term of $(\pp{:init})$. In this
relaxed case, a conjunctive term is visited iff its atomic subterms
are visited, and zero or more of its immediate probabilistic subterms
are visited. The starting state, $\state_0$, for sequential planning
is an {\em abstract} state that captures the unique assumption which
has probability $1$. In $\pp{DORA}$, that is:

@@

In order to make assumptions available to the sequential planner, we
add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
each element, $\prob_i (T_i)$, of each probabilistic term from
$(\pp{:init})$. The physics of these actions is as
follows. $\assumptiveS{i}$ can be executed if no $\assumptiveS{j}$,
$j \neq i$, has been executed from the same probabilistic term, and,
either $(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root conjunct,
or it occurs in $T_k$ for some executed $\assumptiveS{k}$.
%%
Executing $\assumptiveS{i}$ in $\state$ effects a transition to a
successor state $\state^{T_i}$ with probability $\prob_i$, and
$\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
the union of $\state$ with atomic terms from $T_i$ and conflicting
propositions from $s$ deleted. For example, if the action adds
$(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must delete
$(=(\pp{is-in}~\pp{box})~\#)$. State $\state^\bot$ is an added sink,
where $\forall\action$ $\reward(\state^\bot, \action) = 0$
and $\transProb(\state^\bot, \action, \state^\bot) = 1$.

For sequential planning the operators from the POMDP model are
available with semantics that accommodate the above abstraction as
follows. In the deterministic model a proposition $\prop$ can be
thought to have a ternary interpretation at a state \state, as either
{\em true}, written $\prop\in\state$, {\em false}, written
$\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
notation written $\prop\#\in\state$. For example, in $\pp{DORA}$
$(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
$\action\in\actions$ in the deterministic model, if
$\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
or negative effect of $\action$, then $\action$ is not applicable
in \state\ if $\prop\#\in\state$. In other words, an assumptive action
must be applied to make a truth assignment to \prop, and then that
truth assignment could be modified by the DTPDDL domain operators.
%%
Lastly, with regards to switching, we have that if $\action$ has a
precondition that is not true with probability $1$ in the underlying
belief $\bstate$, then immediately before it is executed blindly, the
system switches to a contingent session, discussed in the following
section.

@@

\noindent which is equal to the maximal 
additive contribution a trace of the plan can make to
Eq~\ref{eq:expectedvalue}.  In goal directed problems, where the only
non-zero reward is received at the first transition to a goal state,
the Eq~\ref{eq:tracevalue} criteria gives us the behaviour sought
by \system{FFR$_a$}~\cite{yoon:etal:2007}. From the perspective of
gradient-based reinforcement learning systems, such as William's
REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
under reasonable assumptions the optimal trace identifies a maximal
gradient step from a uniformly random soft-max policy. Finally, it is
worth clarifying that although the above discussion has been in terms
of a trace of length $N$, there is nothing in our approach that
artificially limits the length of candidate sequential plans.
